DDSAnalytics  - Employee Trends

### 2018-04-15 msds6371-group-project-kaggle.Rmd
Update rreport

Model: SalePrice = beta0 + beta1 * GrLivArea + beta2*d1 + beta3*d2 + beta4*d3 + beta5*int1 + beta6*int2 + beta7*int3

@@

Model: SalePrice = beta0 + beta1 * LogGrLivArea + beta3*d2 + beta4*d3 + beta6*centLint2 + beta7*centLint3

@@

#### Models
1. SalePrice    = 19972  +  87.163*GrLivArea   +  68382*d1  + 54705*d2  +  -12416*d3  +  -57.412*int1  +  -32.847*int2  +  31.351*int3
1. LogSalePrice = 5.844  +  0.845*LogGrLivArea +  0.094*d2  + 0.219*d3  +  -0.268*centLint2  +  0.169*centLint3

### Parameters
* Estimates - refer to model #2 above for the parameter estimates
* Interpretation
* Confidence Intervals

@@

### Question 1

#### Plots of Untransformed Data
![Neighborhoods](images/image_001.PNG)*Image 001*
![Untransformed Fit Diagnostics](images/Image_002.PNG)*Image 002*
![Untransformed Proc Reg](images/image_003.PNG)*Image 003*
![Other Neighborhoods](images/Untransformed-Others.png)*Image 004*
![MLR Parameter Estimates](images/Untransformed-MLR-Estimates.png)*Image 005*
![MLR Residual Plots](images/Untransformed-MLR-Residuals.png)*Image 006*

@@

#### Plots of Log-Log Data
![BrkSide](images/Log-Log-BrkSide.png)*Image 007*
![Edwards](images/Log-Log-Edwards.png)*Image 008*
![NAmes](images/Log-Log-NAmes.png)*Image 009*
![Other Neighborhoods](images/Log-Log-Others.png)*Image 010*
![MLR Parameter Estimates](images/Log-Log-MLR-Estimates.png)*Image 011*
![MLR Residual Plots](images/Log-Log-MLR-Residuals.png)*Image 012*

>>>>

```
SalePrice = beta0 + beta1*GrLivArea + beta2*d1 + beta3*d2 + beta4*d3 + beta5*int1 + beta6*int2 + beta7*int3
```

@@

```
SalePrice = beta0 + beta1 * LogGrLivArea + beta3*d2 + beta4*d3 + beta6*centLint2 + beta7*centLint3
```

@@

| Models        | Adusted R^2 | CV Press    |
|---------------|-------------|-------------|
| Untransformed | 0.5696      | 4.00198 E12 |
| Log-Log       | 0.5808      | 97.939      |

Image 007 and Image 008 capture the SAS output for the Untransformned and Log-Log transformed models respectively.

#### Models
1. `SalePrice    = 19972  +  87.163*GrLivArea   +  68382*d1  + 54705*d2  +  -12416*d3  +  -57.412*int1  +  -32.847*int2  +  31.351*int3`
1. `LogSalePrice = 5.844  +  0.845*LogGrLivArea +  0.094*d2  + 0.219*d3  +  -0.268*centLint2  +  0.169*centLint3`

### Parameters
* Estimates - refer to model #2 above for the parameter estimates
* Interpretation
* Confidence Intervals - refer to Image 007 for 95% confidence interval for the individual parameters.

@@

## Question 1 SAS Code

```
/* Data Sets: 
   - Train0 - Kaggle Training Data Set 
   - TrainRed - Kaggle Training Data Set Reduced for question 1  
*/ 
filename CSV URL "https://dl.dropboxusercontent.com/spa/afq05cp80hp4ezn/downloads/public/msds-data/train.csv";

data TRAIN0;
%let _EFIERR_ = 0; /* set the ERROR detection macro variable */
infile CSV
delimiter = ',' MISSOVER DSD lrecl=32767 firstobs=2 ;
/* note regex used : \d\d\d[ ]* */
  informat Id 4. ;
  informat MSSubClass $3. ;
  informat MSZoning $7. ;  /* #31 c (all) */
  informat LotFrontage best32. ;
  informat LotArea best32. ;
  informat Street $4. ;
  informat Alley $4. ;
  informat LotShape $3. ;
  informat LandContour $3. ;
  informat Utilities $6. ;
  informat LotConfig $7. ;
  informat LandSlope $3. ;
  informat Neighborhood $7. ;
  informat Condition1 $6. ;
  informat Condition2 $6. ;
  informat BldgType $6. ;
  informat HouseStyle $6. ;
  informat OverallQual best32. ;
  informat OverallCond best32. ;
  informat YearBuilt best32. ;
  informat YearRemodAdd best32. ;
  informat RoofStyle $7. ;
  informat RoofMatl $7. ;
  informat Exterior1st $7. ;
  informat Exterior2nd $7. ;
  informat MasVnrType $7. ;
  informat MasVnrArea best32. ;
  informat ExterQual $2. ;
  informat ExterCond $2. ;
  informat Foundation $6. ;
  informat BsmtQual $2. ;
  informat BsmtCond $2. ;
  informat BsmtExposure $2. ;
  informat BsmtFinType1 $3. ;
  informat BsmtFinSF1 best32. ;
  informat BsmtFinType2 $3. ;
  informat BsmtFinSF2 best32. ;
  informat BsmtUnfSF best32. ;
  informat TotalBsmtSF best32. ;
  informat Heating $5. ;
  informat HeatingQC $2. ;
  informat CentralAir $1. ;
  informat Electrical $5. ;
  informat _1stFlrSF best32. ;
  informat _2ndFlrSF best32. ;
  informat LowQualFinSF best32. ;
  informat GrLivArea best32. ;
  informat BsmtFullBath best32. ;
  informat BsmtHalfBath best32. ;
  informat FullBath best32. ;
  informat HalfBath best32. ;
  informat BedroomAbvGr best32. ;
  informat KitchenAbvGr best32. ;
  informat KitchenQual $2. ;
  informat TotRmsAbvGrd best32. ;
  informat Functional $4. ;
  informat Fireplaces best32. ;
  informat FireplaceQu $2. ;
  informat GarageType $7. ;
  informat GarageYrBlt best32. ;
  informat GarageFinish $3. ;
  informat GarageCars best32. ;
  informat GarageArea best32. ;
  informat GarageQual $2. ;
  informat GarageCond $2. ;
  informat PavedDrive $1. ;
  informat WoodDeckSF best32. ;
  informat OpenPorchSF best32. ;
  informat EnclosedPorch best32. ;
  informat _3SsnPorch best32. ;
  informat ScreenPorch best32. ;
  informat PoolArea best32. ;
  informat PoolQC $2. ;
  informat Fence $5. ;
  informat MiscFeature $4. ;
  informat MiscVal best32. ;
  informat MoSold best32. ;
  informat YrSold best32. ;
  informat SaleType $5. ;
  informat SaleCondition $7. ;
  informat SalePrice best32. ;
  format Id 4. ;
  format MSSubClass $3. ;
  format MSZoning $2. ;
  format LotFrontage best12. ;
  format LotArea best12. ;
  format Street $4. ;
  format Alley $4. ;
  format LotShape $3. ;
  format LandContour $3. ;
  format Utilities $6. ;
  format LotConfig $7. ;
  format LandSlope $3. ;
  format Neighborhood $7. ;
  format Condition1 $6. ;
  format Condition2 $6. ;
  format BldgType $6. ;
  format HouseStyle $6. ;
  format OverallQual best12. ;
  format OverallCond best12. ;
  format YearBuilt best12. ;
  format YearRemodAdd best12. ;
  format RoofStyle $7. ;
  format RoofMatl $7. ;
  format Exterior1st $7. ;
  format Exterior2nd $7. ;
  format MasVnrType $7. ;
  format MasVnrArea best12. ;
  format ExterQual $2. ;
  format ExterCond $2. ;
  format Foundation $6. ;
  format BsmtQual $2. ;
  format BsmtCond $2. ;
  format BsmtExposure $2. ;
  format BsmtFinType1 $3. ;
  format BsmtFinSF1 best12. ;
  format BsmtFinType2 $3. ;
  format BsmtFinSF2 best12. ;
  format BsmtUnfSF best12. ;
  format TotalBsmtSF best12. ;
  format Heating $5. ;
  format HeatingQC $2. ;
  format CentralAir $1. ;
  format Electrical $5. ;
  format _1stFlrSF best12. ;
  format _2ndFlrSF best12. ;
  format LowQualFinSF best12. ;
  format GrLivArea best12. ;
  format BsmtFullBath best12. ;
  format BsmtHalfBath best12. ;
  format FullBath best12. ;
  format HalfBath best12. ;
  format BedroomAbvGr best12. ;
  format KitchenAbvGr best12. ;
  format KitchenQual $2. ;
  format TotRmsAbvGrd best12. ;
  format Functional $4. ;
  format Fireplaces best12. ;
  format FireplaceQu $2. ;
  format GarageType $7. ;
  format GarageYrBlt best12. ;
  format GarageFinish $3. ;
  format GarageCars best12. ;
  format GarageArea best12. ;
  format GarageQual $2. ;
  format GarageCond $2. ;
  format PavedDrive $1. ;
  format WoodDeckSF best12. ;
  format OpenPorchSF best12. ;
  format EnclosedPorch best12. ;
  format _3SsnPorch best12. ;
  format ScreenPorch best12. ;
  format PoolArea best12. ;
  format PoolQC $2. ;
  format Fence $5. ;
  format MiscFeature $4. ;
  format MiscVal best12. ;
  format MoSold best12. ;
  format YrSold best12. ;
  format SaleType $5. ;
  format SaleCondition $7. ;
  format SalePrice best12. ;
input
  Id
  MSSubClass
  MSZoning $
  LotFrontage $
  LotArea
  Street $
  Alley $
  LotShape $
  LandContour $
  Utilities $
  LotConfig $
  LandSlope $
  Neighborhood $
  Condition1 $
  Condition2 $
  BldgType $
  HouseStyle $
  OverallQual
  OverallCond
  YearBuilt
  YearRemodAdd
  RoofStyle $
  RoofMatl $
  Exterior1st $
  Exterior2nd $
  MasVnrType $
  MasVnrArea $
  ExterQual $
  ExterCond $
  Foundation $
  BsmtQual $
  BsmtCond $
  BsmtExposure $
  BsmtFinType1 $
  BsmtFinSF1
  BsmtFinType2 $
  BsmtFinSF2
  BsmtUnfSF
  TotalBsmtSF
  Heating $
  HeatingQC $
  CentralAir $
  Electrical $
  _1stFlrSF
  _2ndFlrSF
  LowQualFinSF
  GrLivArea
  BsmtFullBath
  BsmtHalfBath
  FullBath
  HalfBath
  BedroomAbvGr
  KitchenAbvGr
  KitchenQual $
  TotRmsAbvGrd
  Functional $
  Fireplaces
  FireplaceQu $
  GarageType $
  GarageYrBlt $
  GarageFinish $
  GarageCars
  GarageArea
  GarageQual $
  GarageCond $
  PavedDrive $
  WoodDeckSF
  OpenPorchSF
  EnclosedPorch
  _3SsnPorch
  ScreenPorch
  PoolArea
  PoolQC $
  Fence $
  MiscFeature $
  MiscVal
  MoSold
  YrSold
  SaleType $
  SaleCondition $
  SalePrice
;
if _ERROR_ then call symputx('_EFIERR_',1);  /* set ERROR detection macro variable */
run;

/* **************************************** */
/* DATA CLEANUP - use proc means to show missing values*/
/* **************************************** */
title "Train0 - Data Set Missing Values";
proc means data=train0 nmiss n; run;

/* fix missing values */
data train0;
  set train0;
  if missing(LotFrontage) then LotFrontage = 0;
run;

title "Train0 - Data Set Fixed Missing Values";
proc means data=train0 nmiss n; run;

title "Train0 - First 10";
proc print data=train0 (obs=10);
run;
title "";
/* **************************************** */
/* END DATA CLEANUP - after second proc means, no missing values */
/* **************************************** */
/********************************/
/** train0 is for Question 2  ***/
/** trainraw3 is Q1 Unmodified **/
/**       / Untransformed Data **/
/********************************/
data train0raw;
  set train0 (keep=Id Neighborhood GrLivArea SalePrice);

  select (Neighborhood);
    when ("NAmes")   Nei=Neighborhood;
    when ("Edwards") Nei=Neighborhood;
    when ("BrkSide") Nei=Neighborhood;
    otherwise Nei="Others";
  end;

  if Nei="Edwards" then d1=1; else d1=0;
  if Nei="NAmes"   then d2=1; else d2=0;
  if Nei="Others"  then d3=1; else d3=0;

  int1 = d1 * GrLivArea; 
  int2 = d2 * GrLivArea; 
  int3 = d3 * GrLivArea;
run;

proc sort data=train0raw;
  by descending GrLivArea;
run;

proc means data=train0raw;
  var GrLivArea d1 d2 d3;
run;

proc sgscatter data=train0raw;
  plot SalePrice*GrLivArea / group=Nei reg;
run;
/**************************************************/
/*************     Model Building      ************/
/* Residual Plot showing possible unequal spread) */
/* Model for Untransformed Data */
proc glm data=train0raw plots=all;
  class Nei(ref="BrkSide");
  model SalePrice=GrLivArea d1 d2 d3 int1 int2 int3 | Nei / solution VIF;
  TITLE "Non Transformed";
run;

proc reg data=train0raw plots=all;
  model SalePrice = GrLivArea d1 d2 d3 int1 int2 int3 / VIF;
  TITLE "Un-Transformed Data";
run;
/******************************************************************************/
/************* Question 1 Tranformed Data *************************************/
/******************************************************************************/
/*************************************************/
/** train0 is for Question 1 - final model data **/
/**        outliers are removed as well         **/
/*************************************************/
data trainRed;
  set train0 (keep=Id Neighborhood GrLivArea SalePrice);

  select (Neighborhood);
    when ("NAmes") Nei=Neighborhood;
    when ("Edwards") Nei="EdwardP";
    when ("BrkSide") Nei="EdwardP";
    otherwise Nei="Others";
  end;

  /* Remove outliers */
  if GrLivArea < 4500;
  /* make sure we do not include missing values */
  if GrLivArea > 0;
  /* make sure we do not include missing values and we dont take log of zero */
  if SalePrice > 0;
    logSales=log(SalePrice);
    logLiv=log(GrLivArea);

  if Nei="NAmes"  then d2=1; else d2=0;
  if Nei="Others" then d3=1; else d3=0;

  l_int2 = d2 * logLiv; 
  l_int3 = d3 * logLiv;

  centLint2=(logLiv - 7.2660236) * (d2 - 0.1543210);
  centLint3=(logLiv - 7.2660236) * (d3 - 0.7386831);
run;

proc sort data=trainRed;
  by descending logLiv;
run;

proc means data=trainRed;
  var logLiv d2 d3;
run;

proc sgscatter data=trainRed;
  plot logSales*logLiv / reg;
  TITLE "Log-Log Transformed Data";
run;

/**************************************************/
/*************     Model Building      ************/

proc glm data=trainRed plots=all;
  class Nei(ref="EdwardP");
  model logSales=logLiv | Nei / solution;
  TITLE "Log-Log Transformed Data";
run;

proc reg data=trainRed plots=all;
  model logSales=logLiv d2 d3 centLint2 centLint3 / VIF;
  TITLE "Log-Log Transformed Data";
run;

/**************************************************/
/*************     Model Comparison   *************/
proc reg data=train0raw plots=all;
  model SalePrice = GrLivArea d1 d2 d3 int1 int2 int3 / PRESS VIF CLM P;
  TITLE "Un-Transformed Data";
run;

proc reg data=trainRed plots=all;
  model logSales=logLiv d2 d3 centLint2 centLint3 / PRESS VIF CLM P;
  TITLE "Log-Log Transformed Data";
run;
```

## Question 1 Plots of Untransformed Data

## Image 001
![Neighborhoods](images/image_001.PNG)

## Image 002
![Fit Diagnostics](images/Image_002.PNG)

## Image 003
![MLR Proc Reg](images/image_003.PNG)

@@

## Question 1 Plots of Log-Log Data

### Image 004
![Neighborhoods](images/image_004.PNG)*Image 004*

### Image 005
![Fit Diagnostics](images/image_005.PNG)*Image 005*

### Image 006
![MLR Proc Reg](images/image_006.PNG)*Image 006*

## Question 1 Miscellaneous Plots

### Image 007
![Untransformed R^2 and PRESS](images/image_007.PNG)

### Image 008
![Log-Log Transformed R^2 and PRESS](images/image_008.PNG)
